---
title: "R Graphics Cookbook"
output: html_notebook
---

### Chap 1: R Basics

```{r}
library(ggplot2)
library(gcookbook)
```

### Chap 2: Quickly Exploring Data

##### Scatterplot

```{r}
# plot(mtcars$wt, mtcars$mpg)

ggplot(mtcars, aes(wt, mpg)) + geom_point()
```

##### Line graph

```{r}
plot(pressure$temperature, pressure$pressure, type ='l')
points(pressure$temperature, pressure$pressure)
```

```{r}
ggplot(pressure, aes(temperature, pressure)) + geom_line()+
  geom_point()
```

##### Bar Graph

```{r}
barplot(BOD$demand, names.arg = BOD$Time)
```

```{r}
barplot(table(mtcars$cyl))
```

```{r}
qplot(factor(mtcars$cyl))
```

```{r}
ggplot(mtcars, aes(factor(cyl))) + geom_bar()
```

```{r}
ggplot(BOD, aes(Time, demand)) + geom_bar(stat = 'identity')
```

##### Histogram

```{r}
hist(mtcars$mpg)
```

```{r}
ggplot(mtcars, aes(mpg)) + geom_histogram(binwidth = 5)
```

##### Boxplots

```{r}
# boxplot(ToothGrowth$supp, ToothGrowth$len)
boxplot(len ~ supp, data = ToothGrowth)
```

```{r}
ggplot(ToothGrowth, aes(supp, len)) + geom_boxplot()
```

##### Plot Function Curve

```{r}
curve(x^3-5*x, from = -4, to=4)
equation <- function(x){
  1/(1 + exp(-x + 10))
}
curve(equation(x), from = 0, to = 20)
curve(equation(x), add = TRUE, col = 'red')

```

```{r}
ggplot(data.frame(x=c(0,20)), aes(x)) + 
  stat_function(fun = equation, geom = 'line')
```

### Chap 3: Bar Graphs

##### Basic Bar Graphs

```{r}
library(gcookbook)

# str(pg_mean)

ggplot(pg_mean, aes(group, weight)) + geom_bar(stat = 'identity', fill = 'lightblue', color = 'black')
```

##### Stacking Bars Together

```{r}
library(gcookbook)

ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar(stat = 'identity')

```

##### Grouping Bars together

```{r}
library(gcookbook)

ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) + 
  geom_bar(stat = 'identity', position = 'dodge', color = 'black') +
  scale_fill_brewer(palette = 'Pastel1')

```

##### Bar Graph of Counts

```{r}
ggplot(diamonds, aes(cut)) + geom_bar(stat = 'count')
```


##### Colors in Bar Graph

```{r}
library(gcookbook)
library(dplyr)
upc <- uspopchange %>% 
  mutate(Change = rank(Change)) %>% 
  filter(Change>40)

ggplot(upc, aes(x = reorder(Abb, Change), y= Change, fill = Region)) + geom_bar(stat = 'identity', color = 'black') +
  scale_fill_manual(values = c("#669933", "#FFCC66"))+
  xlab('State')
  

```


##### 3.5. Colors Negative and Positive Bars

```{r}

library(gcookbook)
library(dplyr)
clim_sub <- climate %>% 
  filter(Source == 'Berkeley', Year >= 1900) %>% 
  mutate(pos = Anomaly10y>=0)

ggplot(clim_sub, aes(Year, Anomaly10y, fill = pos)) + 
  geom_bar(stat = 'identity', position = 'identity', color='black', size =0.25)+
  scale_fill_manual(values = c("#CCEEFF", "#FFDDDD"), guide = FALSE)

```

##### 3.6. Adjusting Bar Width and Spacing

```{r}
library(gcookbook)
ggplot(pg_mean, aes(group, weight)) + geom_bar(stat = 'identity', width = 0.8) 

```

##### 3.7. Making a Stacked Bar Graph

```{r}
library(gcookbook)

ggplot(cabbage_exp, aes(Date, Weight, fill = Cultivar)) +
  geom_bar(stat = 'identity') + 
  guides(fill = guide_legend(reverse = TRUE))
```


##### 3.8. Making a Proportional Stacked Bar Graph

```{r}
library(gcookbook) # For the data set
library(dplyr)
library(plyr)

# cabbage_perc <- cabbage_exp %>% 
#   group_by(Cultivar, Date) %>% 
#   mutate(percent_weight = Weight/sum(Weight) * 100)

cabbage_perc <- ddply(cabbage_exp, 'Date', transform, 
          percent_weight = Weight/sum(Weight) * 100) 

ggplot(cabbage_perc, aes(Date, percent_weight, fill = Cultivar)) + geom_bar(stat = 'identity', color = 'black')+ 
  scale_fill_brewer(palette = 'Pastel1')

```

##### 3.9. Adding Labels to a Bar Graph (see again)

```{r}
library(gcookbook)

ggplot(cabbage_exp, aes(interaction(Date, Cultivar), y = Weight)) +
  geom_bar(stat = 'identity') + 
  geom_text(aes(label = Weight), vjust =1.5, color ='white')

# vjust = -0.2 : text above
```


```{r}
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) + geom_bar(stat = 'identity', position = 'dodge') + geom_text(aes(label = Weight), vjust =1.5, color ='white', position = position_dodge(.9)) 
```


```{r}

# get cummulative sum
cabbage_perc <- ddply(cabbage_perc, 'Date', transform, label_y = cumsum(Weight))

ggplot(cabbage_perc, aes(Date, Weight, fill = Cultivar)) + 
  geom_bar(stat = 'identity') +
  geom_text(aes(y = label_y, label = Weight), vjust = 1.5, color = 'white')
```


##### 3.10. Making a Cleveland Dot Plot

```{r}
library(gcookbook)
tophit_25 <- tophitters2001 %>% 
  select(name, lg, avg) %>% 
  head(20)

ggplot(tophit_25, aes(avg, reorder(name, avg))) +
  geom_point(size = 3) + 
  theme_bw()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(color = 'grey60', linetype = 'dashed')) 

```


```{r}
nameorder <- tophit_25$name[order(tophit_25$lg, tophit_25$avg)]
tophit_25$name <- factor(tophit_25$name, levels=nameorder)

ggplot(tophit_25, aes(avg, name)) + 
  geom_segment(aes(yend = name), xend = 0 , color = 'grey50')+
  geom_point(size = 3, aes(color = lg)) + 
  scale_color_brewer(palette = 'Set1', limits = c('NL', 'AL'), guide = FALSE)+
  theme_bw()+
  theme(panel.grid.major.y = element_blank())+
  facet_grid(lg ~ ., scales = 'free_y' , space = 'free_y')
```

### CHap 4: Line Graphs

##### 4.1. Making a Basic Line Graph

```{r}
ggplot(BOD, aes(Time, demand)) + geom_line()
```

##### 4.2. Adding Points to a Line Graph 

```{r}
ggplot(BOD, aes(Time, demand)) + geom_line() + geom_point()
```

##### 4.3. Making a Line Graph with Multiple Lines

```{r}
# ddply: apply function to column in df
library(plyr)

tooth <- ddply(ToothGrowth, c('supp', 'dose'), summarise, length = mean(len))

ggplot(tooth, aes(x = dose, y= length, fill = supp))+ geom_line() + geom_point(size = 4, shape = 21)

# we can replace fill by color or linetype instead

```

##### 4.4. Changing the Appearance of Lines

```{r}
ggplot(BOD, aes(Time, demand)) + 
  geom_line(linetype = 'dashed', size = 1, color = 'blue')
```

```{r}
library(plyr)

tooth <- ddply(ToothGrowth, c('supp', 'dose'), summarise, length = mean(len))

ggplot(tooth, aes(dose, length, color = supp)) + 
  geom_line(linetype = 'dashed')+
  geom_point(size =3, shape =22, fill = 'white')+
  scale_color_brewer(palette = 'Set1')
```

##### 4.6. Making a Graph with a Shaded Area

```{r}
sunspotyear <- data.frame(
  Year = as.numeric(time(sunspot.year)),
  Sunspot = as.numeric(sunspot.year)
)

ggplot(sunspotyear, aes(Year, Sunspot)) + 
  geom_area(color = 'black', fill = 'green', alpha = 0.6)
```

##### 4.7. Making a Stacked Area Graph

```{r}
library(gcookbook)

ggplot(uspopage, aes(Year, Thousands, fill = AgeGroup))+ 
  geom_area(color = 'black', alpha = 0.4, size = 0.2)+
  scale_fill_brewer(palette = 'Blues', breaks = rev(levels(uspopage$AgeGroup)))
```

##### 4.8. Making a Proportional Stacked Area Graph

```{r}
library(gcookbook)
library(plyr)

uspoppage_perc <- uspopage %>% 
  group_by(Year) %>% 
  mutate(Percent = (Thousands/sum(Thousands)*100))

uspopage_prop <- ddply(uspopage, "Year", transform,
 Percent = Thousands / sum(Thousands) * 100)


ggplot(uspopage_prop, aes(Year, Percent, fill = AgeGroup))+ 
  geom_area(color = 'black', alpha = 0.4)+
  scale_fill_brewer(palette = 'Blues')

```

##### 4.9. Adding a Confidence Region

```{r}
library(gcookbook) 
library(dplyr)

clim_sub <- climate %>% 
  filter(Source == 'Berkeley') %>% 
  select(Year, Anomaly10y, Unc10y)

ggplot(clim_sub, aes(Year, Anomaly10y)) + 
  geom_ribbon(aes(ymin = Anomaly10y-Unc10y, ymax=Anomaly10y+Unc10y), alpha = 0.15)+
  geom_line()
  
# write geom_ribbon first so that it is below the line
```

```{r}
ggplot(clim_sub, aes(Year, Anomaly10y)) + 
  geom_line(aes(y = Anomaly10y-Unc10y), color = 'grey50', linetype = 'dashed') +
  geom_line(aes(y=Anomaly10y+Unc10y), color = 'grey50', linetype = 'dashed')+
  geom_line()
```

### CHAPTER 5: Scatter Plots

##### 5.1. Making a Basic Scatter Plot

```{r}
ggplot(heightweight, aes(ageYear, heightIn)) + 
  geom_point(size = 4, shape = 21)
```


##### 5.2. Grouping Data Points by a Variable Using Shape or Color

```{r}
ggplot(heightweight, aes(ageYear, heightIn, color = sex, shape = sex)) + 
  geom_point()+
  scale_shape_manual(values = c(1,2))+
  scale_color_brewer(palette = 'Set1')

```


##### 5.3. Using Different Point Shapes

##### 5.4. Mapping a Continuous Variable to Color or Size

```{r}
ggplot(heightweight, aes(weightLb, heightIn, fill = ageYear)) +
  geom_point(shape = 21, size = 2.5)+
  scale_fill_gradient(low = 'black', high = 'white', breaks=12:17, guide=guide_legend())
```

```{r}
ggplot(heightweight, aes(ageYear, heightIn, color = sex, size = weightLb))+
  geom_point(alpha= 0.4)+
  scale_size_area()
```

##### 5.5. Dealing with Overplotting

```{r}
ggplot(diamonds, aes(carat, price))+
  stat_bin2d(bins = 50) + 
  scale_fill_gradient(low = 'lightblue' , high = 'red', limits = c(0,60000))

# note: we can import the library(hexbin) instead and use hexbin instead
```

```{r}
ggplot(diamonds, aes(carat, price))+
  geom_jitter(alpha = 0.01)
```


##### 5.6. Adding Fitted Regression Model Lines

```{r}
ggplot(heightweight, aes(ageYear, heightIn)) + 
  geom_point(color = 'grey50')+
  geom_smooth(method = 'lm', se = FALSE, color = 'red')

# note: we can replace 'lm' by 'loess', which is a weighted polynomial function
```

```{r}
ggplot(heightweight, aes(ageYear, heightIn, color = sex)) + 
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)
```

##### 5.7. Adding Fitted Lines from an Existing Model [to do]

```{r}

model <- lm(heightIn ~ ageYear , heightweight)
xmin <- min(heightweight$ageYear)
xmax <- max(heightweight$ageYear)
predicted <- data.frame(ageYear = seq(xmin,xmax, length.out = 100))

predicted$heightIn <- predict(model, predicted)

ggplot(heightweight, aes(ageYear, heightIn))+ 
  geom_point(color = 'grey50')+
  geom_line(data = predicted, size = 1, color ='red')


```

```{r}
library(MASS)

bio <- biopsy %>% 
  mutate(class = if_else(class == 'benign', 0,1))

fitlogistic <- glm(class ~ V1, bio, family = 'binomial')

predictvals <- function(model, xvar, yvar, xrange = NULL, samples = 100 , ...){
  if(is.null(xrange)){
    if(any(class(model) %in% c('lm', 'glm')))
      xrange <- range(model$model[[xvar]])
    else if (any(class(model) %in% 'loess'))
      xrange <- range(model$x)
  }
  
 newdata <- data.frame(x = seq(xrange[1], xrange[2], length.out = samples))
 names(newdata) <- xvar
 newdata[[yvar]] <- predict(model, newdata = newdata, ...)
 newdata
}

glm_predicted <- predictvals(fitlogistic, 'V1', 'class', type = 'response')

ggplot(bio, aes(V1, class, fill = class))+
  geom_jitter(alpha = 0.4, size = 2, shape = 21)+
  geom_line(data = glm_predicted, size = 1)

```


##### 5.8. Adding Fitted Lines from Multiple Existing Models

```{r}
library(plyr)
library(gcookbook)

make_model <- function(data){
  lm(heightIn ~ ageYear , data)
}

# subset data by sex and calculate its regression
models <- dlply(.data = heightweight, .variables =  'sex', .fun = make_model)

pred_vals <- ldply(models, .fun = , predictvals, xvar="ageYear", yvar="heightIn")

ggplot(heightweight, aes(ageYear, heightIn, color = sex)) +
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_line(data = pred_vals, color = 'black')

```


##### 5.9. Adding Annotations with Model Coefficients

```{r}
ggplot(heightweight, aes(ageYear, heightIn, color = sex)) +
  geom_point()+
  geom_smooth(method = 'lm')+
  geom_line(data = pred_vals, color = 'black')+
  annotate('text', label = 'r^2==0.42', x=16.5, y=52, parse=TRUE)

# note: parse=TRUE makes string read as maths formula
```

##### 5.10. Adding Marginal Rugs to a Scatter Plot

```{r}

ggplot(faithful, aes(eruptions, waiting))+
  geom_point()+
  geom_rug(position = 'jitter', size = 0.2)

```

##### 5.11. Labeling Points in a Scatter Plot

```{r}

library(dplyr)

countries_sub <- countries %>% 
  filter(Year == 2009, healthexp >= 2000)

sp <- ggplot(countries_sub, aes(healthexp, infmortality))+
  geom_point()

sp +
  annotate('text', x = 4350, y = 5.4 , label = 'Canada')+
  annotate("text", x=7400, y=6.8, label="USA")

sp +
  geom_text(aes(label = Name), size = 4)

sp +
  geom_label(aes(label = Name), size = 4)

```


##### 5.12. Creating a Balloon Plot

```{r}
library(dplyr)

countries_sub <- countries %>% 
  filter(Year == 2009, 
         Name %in%  c("Canada", "Ireland", "United Kingdom", "United States",
 "New Zealand", "Iceland", "Japan", "Luxembourg",
 "Netherlands", "Switzerland"))

ggplot(countries_sub, aes(healthexp, infmortality, size =GDP))+
  geom_point(shape = 21, color = 'black', fill = 'cornsilk') + 
  scale_size_area(max_size = 15) # make circle bigger


```

```{r}
library(reshape2)

hec <- HairEyeColor[,,'Male'] + HairEyeColor[,,'Female']
hec <-  melt(hec, value.name = 'count')
  
ggplot(hec, aes(Eye, Hair)) +
  geom_point(aes(size = count), color = 'black', fill = 'cornsilk', shape = 21) +
  scale_size_area(max_size = 10)+
  geom_text(aes(label = count), vjust = -1, color = 'grey60')

```



##### 5.13. Making a Scatter Plot Matrix

```{r}
library(dplyr)

countries_sub <- countries %>% 
  filter(Year == 2009) %>% 
  dplyr:: select(Name, GDP, laborrate, healthexp, infmortality)

pairs(countries_sub)

```

```{r}
## custom 

```


### Chap 6: Summarized Data Distributions

```{r}
ggplot(faithful, aes(waiting))+ 
  geom_histogram(binwidth = 5, fill = 'white', color = 'black')

# note: add paramter origin = 30 to start at 30
```

##### 6.1. Making a Basic Histogram

```{r}
binsize <- diff(range(faithful$waiting)/15)
ggplot(diamonds, aes(carat))+ 
  geom_histogram(binsize = binsize, color= 'black', fill = 'white')
```



##### 6.2. Making Multiple Histograms from Grouped Data

```{r}
library(MASS)
library(dplyr)

birth_2 <- birthwt %>% 
  mutate(smoke = ifelse( smoke ==1, 'Smokes', 'No Smokes'))

ggplot(birth_2, aes(bwt))+
  geom_histogram(fill = 'white', color = 'black')+
  facet_grid(smoke ~ . )
```

##### 6.3. Making a Density Curve

```{r}
ggplot(faithful, aes(x=waiting)) + 
  geom_density(stat = 'density', color = 'red')+
  expand_limits(y=0)
```

```{r}
ggplot(faithful, aes( waiting)) +
  geom_line(stat = 'density', adjust = 0.25, color = 'red')+
  geom_line(stat = 'density', adjust = 0.5, color = 'blue')+
  geom_line(stat = 'density')
```

```{r}
ggplot(faithful, aes(waiting)) +
  geom_density(fill = 'blue', alpha =.2, color = NA)+
  geom_line(stat = 'density')
```

```{r}
ggplot(faithful, aes(waiting, y = ..density..))+
  geom_histogram(color = 'grey85', size = 0.2, fill = 'cornsilk')+
  geom_density()+
  xlim(35, 105)

```

##### 6.4. Making Multiple Density Curves from Grouped Data

```{r}
library(MASS)
library(dplyr)

birth_2 <- birthwt %>% 
  mutate(smoke = as.factor(smoke))

ggplot(birth_2, aes(age, fill = smoke))+
  geom_density(alpha = 0.4, size = 0.8)
```


```{r}
ggplot(birth_2, aes(age, y = ..density..))+
  geom_histogram(bins = 15, fill = 'cornsilk')+
  geom_density()+
  facet_grid(smoke~ . )
```


##### 6.5. Making a Frequency Polygon

```{r}
binsize <- diff(range(faithful$waiting))/15
ggplot(faithful, aes(waiting))+ 
  geom_freqpoly(binwidth = binsize)
```

##### 6.6. Making a Basic Box Plot

```{r}
library(MASS)

ggplot(birthwt, aes(x=factor(race), y = lwt)) +
  geom_boxplot(width = .5, outlier.size = 1.5, outlier.shape = 21)
```

```{r}
ggplot(birthwt, aes(x=1, lwt))+
  geom_boxplot()+
  scale_x_continuous(breaks = NULL)+
  theme(axis.title.x = element_blank())
```

##### 6.7. Adding Notches to a Box Plot

```{r}
library(MASS)

ggplot(birthwt, aes(x = factor(race), bwt))+
  geom_boxplot(notch = TRUE)

# note: notches are used to see if median differ
```

##### 6.8. Adding Means to a Box Plot

```{r}
library(MASS)

ggplot(birthwt, aes(x = factor(race), bwt))+
  geom_boxplot()+
  stat_summary(fun.y = 'mean', geom = 'point', shape = 23, size = 3 , fill ='white')

```


##### 6.9. Making a Violin Plot

```{r}
ggplot(heightweight, aes(x= sex, y= heightIn))+
  geom_violin(trim = TRUE, scale = 'count')+
  geom_boxplot(width = .1, fill = 'black', outlier.colour = NA)+
  stat_summary(fun.y = median, geom = 'point', fill = 'white', shape =21, size =2.5)
```

```{r}
ggplot(heightweight, aes(sex, heightIn))+
  geom_violin(adjust =.5)
```

##### 6.10. Making a Dot Plot

```{r}
ggplot(countries_sub, aes(x=infmortality))+
  geom_dotplot()+
  geom_rug()+
  scale_y_continuous(breaks = NULL)+
  theme(axis.title.y = element_blank())
```

```{r}
ggplot(countries_sub, aes(x=infmortality))+
  geom_dotplot(method = 'histodot', binwidth = 2.5, stackdir = 'centerwhole')+
  geom_rug()+
  scale_y_continuous(breaks = NULL)+
  theme(axis.title.y = element_blank())
```


##### 6.11. Making Multiple Dot Plots for Grouped Data

```{r}
ggplot(heightweight, aes(x=sex, y = heightIn))+
  geom_dotplot(stackdir = 'center', binaxis = 'y', binwidth = .5, fill = NA)
```

```{r}
ggplot(heightweight, aes(x=sex, y=heightIn)) +
 geom_boxplot(aes(x=as.numeric(sex) + .2, group=sex), width=.25) +
 geom_dotplot(aes(x=as.numeric(sex) - .2, group=sex), binaxis="y",
 binwidth=.5, stackdir="center") +
 scale_x_continuous(breaks=1:nlevels(heightweight$sex),
 labels=levels(heightweight$sex))
```



##### 6.12. Making a Density Plot of Two-Dimensional Data

```{r}
ggplot(faithful, aes(eruptions, waiting))+
  geom_point()+
  stat_density2d(aes(color = ..level..))
```

```{r}
ggplot(faithful, aes(eruptions, waiting))+
  stat_density2d(aes(fill = ..density..), geom = 'raster', contour=FALSE)
```


### Chap 7: Annotations

##### 7.1. Adding Text Annotations

```{r}
p <- ggplot(faithful, aes(eruptions, waiting))+
  geom_point()

p + annotate('text', x = 3, y =48 , label = 'Group 1', family = 'serif', fontface = 'italic', color = 'darkred', size =3)+
  annotate('text', x = 4.5, y =66, label = 'Group 2 ')
  
```


##### 7.2. Using Mathematical Expressions in Annotations

```{r}
p <- ggplot(data.frame(x=c (-3,3)), aes(x)) + stat_function(fun = dnorm)

p + annotate('text', x =2, y= 3, parse = TRUE,
             label = 'frac(1, sqrt(2*pi)) * e ^ {-x^2/2}')

```

##### 7.3. Adding Lines

```{r}
p <- ggplot(heightweight, aes(ageYear, heightIn, color = sex))+
  geom_point()

p+
  geom_hline(yintercept = mean(heightweight$heightIn))+
  geom_vline(xintercept = mean(heightweight$ageYear))
```

```{r}
library(dplyr)
library(gcookbook)

hw_means_by_sex <- heightweight %>% 
  group_by(sex) %>% 
  summarise(heightIn_mean = mean(heightIn), n = n())

ggplot(heightweight, aes(ageYear, heightIn, color =sex)) +
  geom_point()+
 geom_hline(data=hw_means_by_sex, yintercept = hw_means_by_sex$heightIn_mean)
```

```{r}
library(plyr)

hw_means <- ddply(heightweight, 'sex', summarise, heightIn = mean(heightIn))

p +
  geom_hline(data=hw_means, aes(yintercept = heightIn, color = sex ), linetype = 'dotted', size =1)

```


##### 7.4. Adding Line Segments and Arrows

```{r}
climate_sub <- climate %>%
  filter(Source == 'Berkeley')

p <- ggplot(climate_sub, aes(Year, Anomaly10y))+
  geom_line()

p +
  annotate('segment', x = 1900, xend = 1975, y= - .5, yend= - .5)+
  annotate('segment', x = 1850, xend = 1825, y = -0.85, yend = -.95, arrow = arrow(), size =2, color = 'red') + 
  annotate('segment', x = 1875, xend = 1900 , y = 0.5, yend = 0.5, arrow = arrow(ends = 'both', angle =90, length = unit(.2, 'cm')))
  

```


##### 7.5. Adding a Shaded Rectangle

```{r}
p + 
  annotate('rect', xmin = 1950, xmax = 1975, ymin = -1, ymax = 1, alpha=0.1, fill = 'black')
```


##### 7.6. Highlighting an Item

```{r}
plant_2 <- PlantGrowth %>% 
  mutate(hl = if_else(group == 'trt2', 'yes', 'no'))

ggplot(plant_2, aes(group, weight, fill = hl)) + 
  geom_boxplot()+
  scale_fill_manual(values = c("grey85", "#FFDDCC"), guide = FALSE)

```

##### 7.7. Adding Error Bars

```{r}

cabbage_2 <- cabbage_exp %>% 
  filter(Cultivar == 'c39')

p <- ggplot(cabbage_2, aes(x= Date, y=Weight))

p + 
  geom_bar(fill = 'white', color = 'black', stat = 'identity')+
  geom_errorbar(aes(ymin= Weight -se, ymax = Weight +se), width=.2)

```
```{r}
p + 
  geom_line(aes(group=1))+
  geom_point(size =3)+
  geom_errorbar(aes(ymin = Weight - se, ymax = Weight + se))
```

```{r}
ggplot(cabbage_exp, aes(x= Date, y=Weight, fill = Cultivar))+
  geom_bar(stat = 'identity', position = 'dodge', color = 'black')+
  geom_errorbar(aes(ymin = Weight -  se, ymax = Weight +se), position = position_dodge(0.9), width = 0.2)
```

##### 7.8. Adding Annotations to Individual Facets

```{r}

p_labels <- data.frame(drv = c('4', 'f', 'r'), label = c('4dw', 'Front', 'Rear'))

ggplot(mpg, aes(displ, hwy))+
  geom_point()+
  facet_grid(.~ drv)+
  geom_text(aes(label = label), data = p_labels, x = 6, y = 40)
```

```{r}
# failed
lm_labels <-  function(data){
  model <- lm(hwy ~ displ, data = data)
  formula <- sprintf("italic(y) == %.2f %+.2f * italic(x)", round(coef(model)[1], 2), round(coef(model)[2], 2))
  r <- cor(data$displ, data$hwy)
  r2 <- sprintf('italic(R^2) == %.2f', r^2)
  data.frame(formula = formula , r2=r2, stringsAsFactors = FALSE)
}

labels <- ddply(mpg, 'drv', lm_labels)

p + geom_smooth(method = 'lm', se=FALSE) +
  geom_text(x=3,y=40, aes(label = formula), data = labels, parse = TRUE, hjust =0)+
  geom_text(x=3, y=35, aes(label=r2), data=labels, parse=TRUE, hjust=0)

```


### Chap 8: Axis

##### 8.1. Swapping X- and Y-Axes

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight))+
  geom_boxplot()+
  coord_flip()
```

##### 8.2. Setting the Range of a Continuous Axis

```{r}
library(ggplot2)
ggplot(PlantGrowth, aes(group, weight)) +
  geom_boxplot() + 
  ylim(0, max(PlantGrowth$weight)) +
  coord_flip()+
  scale_y_continuous(breaks = NULL)
```

##### 8.3. Reversing a Continuous Axis

```{r}
ggplot(PlantGrowth, aes(group, weight))+
  geom_boxplot()+
  scale_y_reverse() 
```


##### 8.4. Changing the Order of Items on a Categorical Axis

```{r}
ggplot(PlantGrowth, aes(group, weight))+
  geom_boxplot() + 
  scale_x_discrete(limits = c('trt2', 'trt1', 'ctrl'))
```

```{r}
ggplot(PlantGrowth, aes(group, weight)) + 
  geom_boxplot() + 
  scale_x_discrete(limits = rev(levels(PlantGrowth$group)))
```

##### 8.5. Setting the Scaling Ratio of the X- and Y-Axes

```{r}
library(gcookbook)
ggplot(marathon, aes(Half, Full)) + 
  geom_point()+ 
  coord_fixed(ratio = 1/2)
```


##### 8.6. Setting the Positions of Tick Marks

```{r}
ggplot(PlantGrowth, aes(group, weight)) + 
  geom_boxplot() + 
  scale_y_continuous(breaks = c(4, 4.25, 4.5, 5, 6, 8))
```


##### 8.7. Removing Tick Marks and Labels

```{r}
ggplot(PlantGrowth, aes(group, weight)) + 
  geom_boxplot()+
  theme(axis.text.y = element_blank(), 
        axis.ticks = element_blank())
```


##### 8.8. Changing the Text of Tick Labels

```{r}
ggplot(heightweight, aes(ageYear, heightIn)) + 
  geom_point()+
  scale_y_continuous(breaks = c(50,56,60,66,72),
  labels = c('tiny', 'short', 'small', 'medium', 'tall'))
```

```{r}
## change ytext to feet
```


##### 8.9. Changing the Appearance of Tick Labels

```{r}
ggplot(PlantGrowth, aes(group, weight)) + 
  geom_boxplot() + 
  scale_x_discrete(breaks = c('ctrl', 'trt1', 'trt2'), labels = c('Control', 'Treatment 1', 'Treatment 2'))+
  theme(axis.text.x = element_text(angle = 45, hjust =1, family = 'Times', face = 'italic'))
```

##### 8.10. Changing the Text of Axis Labels

```{r}
ggplot(PlantGrowth, aes(group, weight)) + 
  geom_point()+
  xlab('Test Groups')+
  ylab('weight in kg')
```

##### 8.11. Removing Axis Labels

```{r}
ggplot(PlantGrowth, aes(group, weight))+
  geom_point()+
  theme(axis.title = element_blank())
```

##### 8.12. Changing the Appearance of Axis Labels

```{r}
ggplot(PlantGrowth, aes(group, weight))+
  geom_point()+
  theme(axis.title.x = element_text(face = 'italic', colour = 'blue'))
```

##### 8.13. Showing Lines Along the Axes

```{r}
ggplot(PlantGrowth, aes(group, weight))+
  geom_point()+
  theme(axis.line = element_line(colour = 'black', size = 1))
```

##### 8.14. Using a Logarithmic Axis

```{r}
library(MASS)
ggplot(Animals, aes(body, brain)) + 
  geom_text(aes(label = rownames(Animals)))+
  scale_x_log10()+
  scale_y_log10()
```

##### 8.15. Adding Ticks for a Logarithmic Axis

```{r}
library(scales)
ggplot(Animals, aes(body, brain)) + 
  geom_text(aes(label = rownames(Animals)))

## todo
```

##### 8.16. Making a Circular Graph

```{r}
ggplot(wind, aes(DirCat, fill = SpeedCat))+
  geom_histogram( binwidth = 15, boundary = -7.5, color = 'black', size = .25)+
  guides(fill = guide_legend(reverse = TRUE))+
  coord_polar()+
  scale_fill_brewer(palette = 'Blues')
```

##### 8.17. Using Dates on an Axis

```{r}
str(economics)
library(dplyr)

econ <- economics %>% 
  mutate(date >= as.Date('1992-01-05'), date< as.Date('1993-06-30'))

p <- ggplot(econ, aes(date, psavert))+
  geom_line()

date_breaks <- seq(as.Date("1992-06-01"), as.Date("1993-06-01"), by = '2 month')

p +
  scale_x_date(breaks = date_breaks)+
  theme(axis.text.x = element_text(angle = 30, hjust =1))

```

##### 8.18. Using Relative Times on an Axis

```{r}

```

### Chap 9: Controlling the Overall Appearance of Graphs

##### 9.1. Setting the Title of a Graph

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

p + 
  ggtitle('Children\'s Age and Height ') + 
  theme(plot.title = element_text(hjust = 0.5))
```


##### 9.2. Changing the Appearance of Text

```{r}

```

##### 9.3. Using Themes

```{r}

```

##### 9.4. Changing the Appearance of Theme Elements
```{r}

```

##### 9.5. Creating Your Own Themes
```{r}
custom_theme <- theme_bw() + 
  theme(text = element_text(colour = 'blue'),
        axis.title = element_text(size = 2, hjust = 0.5))

p + 
  custom_theme
```


##### 9.6. Hiding Grid Lines
```{r}
p + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```


### Chap 10: Legends

##### 10.1. Removing the Legend
```{r}
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()

p + 
  guides(fill = FALSE)

p + 
  theme(legend.position = 'none')

```

##### 10.2. Changing the Position of a Legend
```{r}
p <- ggplot(PlantGrowth, aes(group, weight, fill = group)) + 
  geom_boxplot()+
  scale_fill_brewer(palette = 'Pastel2')

p + 
  theme(legend.position = 'top',
        legend.background = element_rect(fill = 'white', colour = 'black')) 
```

##### 10.3. Changing the Order of Items in a Legend
```{r}
p+
  scale_fill_discrete(limits = c('trt1', 'trt2', 'ctrl'))

p + 
  scale_fill_brewer(limits = c('trt1', 'trt2', 'ctrl'), palette = 'Pastel2')

```

##### 10.4. Reversing the Order of Items in a Legend
```{r}
p + 
  guides(fill = guide_legend(reverse =TRUE))
```

##### 10.5. Changing a Legend Title
```{r}

```

##### 10.6. Changing the Appearance of a Legend Title

```{r}

```
##### 10.7. Removing a Legend Title
```{r}

```

##### 10.8. Changing the Labels in a Legend
```{r}

```

##### 10.9. Changing the Appearance of Legend Labels
```{r}

```

##### 10.10. Using Labels with Multiple Lines of Text
```{r}

```

### Chap 11: Facets

##### 11.1. Splitting Data into Subplots with Facets
```{r}

```

##### 11.2. Using Facets with Different Axes
```{r}

```

##### 11.3. Changing the Text of Facet Labels
```{r}

```

##### 11.4. Changing the Appearance of Facet Labels and Headers
```{r}

```



### Chap 12: Using Colors in Plots
```{r}

```

##### 12.1. Setting the Colors of Objects
```{r}

```

##### 12.2. Mapping Variables to Colors
```{r}

```

##### 12.3. Using a Different Palette for a Discrete Variable
```{r}

```

##### 12.5. Using a Colorblind-Friendly Palette
```{r}

```

##### 12.6. Using a Manually Defined Palette for a Continuous Variable
```{r}

```

##### 12.7. Coloring a Shaded Region Based on Value

```{r}

```

### Chap 13: Miscellaneous Graphs

##### 13.1. Making a Correlation Matrix
```{r}

```

##### 13.2. Plotting a Function
```{r}

```

##### 13.3. Shading a Subregion Under a Function Curve
```{r}

```

##### 13.4. Creating a Network Graph
```{r}

```

##### 13.5. Using Text Labels in a Network Graph
```{r}

```

##### 13.6. Creating a Heat Map
```{r}

```

##### 13.7. Creating a Three-Dimensional Scatter Plot
```{r}

```

##### 13.8. Adding a Prediction Surface to a ThreeDimensional Plot
```{r}

```

##### 13.9. Saving a Three-Dimensional Plot
```{r}

```

##### 13.10. Animating a Three-Dimensional Plot
```{r}

```

##### 13.11. Creating a Dendrogram
```{r}

```

##### 13.12. Creating a Vector Field
```{r}

```

##### 13.13. Creating a QQ Plot
```{r}

```

##### 13.14. Creating a Graph of an Empirical Cumulative Distribution Function

```{r}

```

##### 13.15. Creating a Mosaic Plot
```{r}

```

##### 13.16. Creating a Pie Chart
```{r}

```

##### 13.17. Creating a Map
```{r}

```

##### 13.18. Creating a Choropleth Map
```{r}

```

##### 13.19. Making a Map with a Clean Background
```{r}

```

##### 13.20. Creating a Map from a Shapefile
```{r}

```


### Chap 14: Output for Presentation


##### 14.1. Outputting to PDF Vector Files
```{r}

```

##### 14.2. Outputting to SVG Vector Files
```{r}

```

##### 14.3. Outputting to WMF Vector Files
```{r}

```

##### 14.4. Editing a Vector Output File
```{r}

```

##### 14.5. Outputting to Bitmap (PNG/TIFF) Files
```{r}

```

##### 14.6. Using Fonts in PDF Files
```{r}

```

##### 14.7. Using Fonts in Windows Bitmap or Screen Output
```{r}

```

### Chap 15: Getting Your Data into Shape


##### 15.1. Creating a Data Frame
```{r}

```

##### 15.2. Getting Information About a Data Structure
```{r}

```

##### 15.3. Adding a Column to a Data Frame
```{r}

```



##### 15.4. Deleting a Column from a Data Frame
```{r}

```

##### 15.5. Renaming Columns in a Data Frame
```{r}

```

##### 15.6. Reordering Columns in a Data Frame
```{r}

```

##### 15.7. Getting a Subset of a Data Frame
```{r}

```

##### 15.8. Changing the Order of Factor Levels
```{r}

```

##### 15.9. Changing the Order of Factor Levels Based on Data Values
```{r}

```

##### 15.10. Changing the Names of Factor Levels
```{r}

```


##### 15.11. Removing Unused Levels from a Factor
```{r}

```

##### 15.12. Changing the Names of Items in a Character Vector
```{r}

```

##### 15.13. Recoding a Categorical Variable to Another Categorical Variable
```{r}

```



##### 15.14. Recoding a Continuous Variable to a Categorical Variable

```{r}

```

##### 15.15. Transforming Variables
```{r}

```

##### 15.16. Transforming Variables by Group
```{r}

```

##### 15.17. Summarizing Data by Groups
```{r}

```

##### 15.18. Summarizing Data with Standard Errors and Confidence Intervals
```{r}

```



##### 15.19. Converting Data from Wide to Long
```{r}

```

##### 15.20. Converting Data from Long to Wide
```{r}

```

##### 15.21. Converting a Time Series Object to Times and Values

```{r}

```



